---
title: "6-MissingData"
output:
  pdf_document:
    number_sections: yes
    df_print: tibble
    latex_engine: xelatex
---

```{r setup6, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, out.width="50%")
# Set root directory from Rmd dir to project dir
opts_knit$set(root.dir = "..")
```

```{r libraries6, include=FALSE}
source("Scripts/0-setup.R")
```

```{r data6, include=FALSE}
dat <- readRDS("Data/prepare2_dat.rds")
```

Description of missing data analysis.

## Assess missingness

### Visualise data frame

A graphical representation of the data types and proportion of missing values
for each variable is shown below. Ancillary variables that are not required for
modelling are not shown. Variables in the data frame are plotted on the x-axis
and each observation (i.e. participant) is plotted on the y-axis. Missing values
are shaded grey.

```{r visualise}
# Ancillary variables (non-predictor, non-outcome)
# Unimportant if there are missing values for these
anc_vars <- c(
  "adm_session",
  "adm_uid",
  "adm_datetime",
  "pi_sex",
  "pi_age",
  "pi_type",
  "pi_admreason",
  "oh_delivery",
  "dis_session",
  "dis_uid",
  "outcome",
  "outcome_datetime",
  "adm_dur",
  "time",
  "diagnosis",
  "diagnosis_other",
  "diagnosis2",
  "diagnosis2_other",
  "cause_death",
  "cause_death_other",
  "cause_death2",
  "cause_death2_other"
  )

# Candidate predictor and outcome variables (model variables)
model_vars <- names(dat)[!names(dat) %in% anc_vars]

# Data frames
anc_df <- dat %>% select(all_of(anc_vars))
model_df <- dat %>% select(all_of(model_vars))

vis_dat(model_df)
```

### Variable-wise missingness

The number and percentage of missing values for each variable is shown below.
In total, `r sum(miss_var_summary(model_df)$pct_miss > 0)` variables had
missing values.

```{r vmiss}
model_df %>% miss_var_summary() %>% print(n = Inf)
```

### Case-wise missingness

Most participants had no missing data and, among those who did, the majority
were only missing values for one predictor (most commonly temperature at
admission).

```{r na_patterns}
model_df %>% miss_case_table() %>% print(n = Inf)

aggr_plot <- model_df %>%
  select(which(colMeans(is.na(.)) > 0)) %>%
  VIM::aggr(plot = FALSE)

plot(
  aggr_plot,
  sortVars = T,
  col=c("grey", mdc(2)),
  numbers = TRUE,
  prop = c(TRUE, FALSE),
  cex.axis = .8,
  gap = 3,
  oma = c(6,5,1,2))

na_patterns <- Hmisc::naclus(model_df)
Hmisc::naplot(na_patterns, which=c("na per obs"), col = "red")
```

The dendrogram below shows predictors that were commonly missing together.

```{r na_patterns2}
plot(na_patterns, ylab = "Fraction of NAs in common", col = "red")
```

### Relationship between missing temperature and the study outcome

There was no evidence of an association between having a missing value for
temperature at admission and the primary outcome of early-onset sepsis:

```{r na_outcome, message=FALSE}
dat <- dat %>%
  mutate(na_temp = if_else(is.na(et_temp), 1, 0))

table(dat$na_temp, dat$sepsis)

na_mod1 <- glm(sepsis ~ na_temp, data = dat, family = "binomial")
summary(na_mod1)
tbl_regression(na_mod1, exponentiate = TRUE)
```

### Relationship between missing temperature and time

Towards the start of the Neotree project, there was a limited number of
thermometers available to measure temperature and, therefore, time since the
start of the study is a plausible predictor of missingness.

Indeed, most missing values for temperature at admission occurred near the start
of data collection. This suggests that temperature was missing at random (MAR)
conditional on time since start of the project.

The matrix plot below shows missing values in red, with each participant sorted
by their admission date (i.e. time since the start of data collection).

```{r na_time}
dat %>%
  select(all_of(model_vars), time) %>%
  select(which(colMeans(is.na(.)) > 0), sepsis, time) %>%
  VIM::matrixplot(sortby = "time")
```

Furthermore, the below figure and a logistic regression analysis demonstrate
that time since the start of data collection was a significant predictor of
temperature at admission being missing.

Notably, the average recorded temperature was approximately 0.5°C higher during
the first 100 days compared to the rest of the data collection period. It is
plausible that, during the first 100 days, healthcare workers were more likely
to record temperature for ‘sicker’ babies who were thus more likely to have an
elevated temperature. Nevertheless, a wide range of participant characteristics
were collected by the Neotree app and were included in the imputation model.

```{r na_time2, message=FALSE}
dat %>%
  ggplot(aes(x = time, y = et_temp)) +
  geom_miss_point() +
  scale_y_continuous(breaks = seq(30, 40, 1)) +
  scale_colour_manual(
    values = c(mdc(2), "grey"),
    name = "Missingness",
    labels = c("Missing", "Not missing")
    ) +
  geom_smooth(colour = "grey35") +
  labs(x = "time (days)") +
  theme_classic2() +
  theme(legend.position = "bottom") +
  NULL

na_mod2 <- glm(na_temp ~ time, data = dat, family = "binomial")
summary(na_mod2)
tbl_regression(na_mod2, exponentiate = TRUE)
```

## Impute missing values

The imputation model contained all candidate predictors, the outcome of sepsis,
and ancillary variables included in the descriptive analysis or that were
determined to predict missingness (i.e. time, see above).

```{r impute, include=FALSE}
# Define objects for imputation model --------------------

# Final set of candidate predictors
predictors <- c(
  "pi_gest",
  "et_bw",
  "oh_matfever",
  "oh_offliquor",
  "co_prom",
  "et_grunt",
  "et_rr",
  "et_hr",
  "et_temp",
  "oe_activity",
  "oe_nasalflare",
  "oe_retractions",
  "oe_grunt",
  "oe_wob"
)

# Excluded predictors
excluded <- c(
  "et_cyanosis",
  "et_seizures",
  "oe_fontanelle",
  "oe_colour",
  "oe_abdodist",
  "oe_omphalitis",
  "oe_abskin",
  "hx_vomit"
)

# Final candidate predictors with missing values (i.e., to impute)
impute_vars <- c(
  "et_temp",
  "et_bw",
  "oe_wob",
  "et_rr",
  "et_hr",
  "et_grunt",
  "oe_nasalflare",
  "oe_retractions",
  "oe_grunt",
  "oh_matfever",
  "oh_offliquor"
)

# All variables (i.e., to inform the imputation model)
inform_vars <- c(
  predictors,
  excluded,
  "sepsis",
  "time",
  "pi_sex",
  "pi_age",
  "oh_delivery",
  "outcome"
)

ivars_df <- dat %>% select(all_of(inform_vars))

# Set MICE parameters --------------------

# Dry run to get defaults
set.seed(123)

dry_imp <- mice(
  ivars_df,
  m = 1,
  maxit = 0,
  defaultMethod = c("pmm", "logreg", "polyreg", "polr")
)

# Get predictor matrix
pm <- dry_imp$predictorMatrix

# Get methods
method <- dry_imp$method
method
```

Data were assumed to be MAR and 40 imputed datasets were created with 20
iterations. There is no consensus on the optimal number of imputations for
multiple imputation, but 40 was chosen based on
`r round(sum(miss_case_table(model_df)$pct_cases[miss_case_table(model_df)$n_miss_in_case >= 1]), 1)`%
of participants having at least one missing value.

```{r impute2, include=FALSE}
# Impute --------------------
# m = 40, iterations = 20
set.seed(123)

# Assign output of "5-impute.R" so no need to recompute here...

# imp <- mice(
#   ivars_df,
#   m = 40,
#   maxit = 20,
#   predictorMatrix = pm,
#   method = method
# )

imp <- readRDS("Data/imputed_dfs.rds")

```

The performance of the imputation model is shown below:

```{r impute3}
print("Imputation method for each variable...")
imp$method

# Diagnostic plots
print("Diagnostic plots...")

order <- c("et_temp", "et_bw", "et_rr", "et_hr")
plot(imp, y = order, layout = c(2,4))
densityplot(imp, ~ et_temp + et_bw + et_rr + et_hr)
stripplot(
  imp,
  et_temp + et_bw + et_rr + et_hr ~ .imp,
  col= c("grey", mdc(2)),
  pch=c(1,20),
  cex = 1,
  layout = c(2,2)
  )
```

