---
title: "7-DescriptiveStatistics"
output:
  pdf_document:
    number_sections: yes
    df_print: tibble
    latex_engine: xelatex
---

```{r setup7, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, out.width="50%")
# Set root directory from Rmd dir to project dir
opts_knit$set(root.dir = "..")
```

```{r libraries7, include=FALSE}
source("Scripts/0-setup.R")
```

```{r data7, include=FALSE}
dat <- readRDS("Data/prepare2_dat.rds")
```

Descriptive analysis of included participants. Data are presented for the
observed data only (i.e. before MICE) using pairwise deletion of missing values.

## Distribution of continuous variables

```{r distributions}
t1_num <- c("pi_gest", "et_bw", "adm_dur")
labels <- c("gestation (weeks)", "birth weight (grams)", "admission duration (days)")

# All participants
dat %>%
  select(all_of(t1_num)) %>%
  map2(labels, ~(
    ggplot(data = dat, aes(x = .)) +
      geom_histogram(bins = 12, fill = "grey30", colour = "black") +
      labs(x = .y) +
      theme_classic2() +
      NULL
    )
)

# Facet wrapped by sepsis vs. no sepsis
dat %>%
  select(all_of(t1_num)) %>%
  map2(labels, ~(
    ggplot(data = dat, aes(x = ., fill = dat$sepsis)) +
      geom_histogram(bins = 12, colour = "black") +
      scale_fill_brewer() +
      labs(x = .y) +
      theme_classic2() +
      theme(legend.position = "none") +
      facet_wrap(~ dat$sepsis, nrow = 2) +
      NULL
    )
)

```

"Gestation" and "birth weight" are approximately normally distributed, while
"admission duration" is very right-skewed.

## Table 1

Table summarising the characteristics of included participants:

```{r table1}
t1_vars <- c(
  "pi_sex",
  "pi_gest",
  "et_bw",
  "pi_age",
  "pi_type",
  "oh_delivery",
  "adm_dur",
  "outcome",
  "sepsis"
  )

labels <- list(
  pi_gest ~ "Gestational age",
  et_bw ~ "Birth weight",
  adm_dur ~ "Admission duration",
  pi_sex ~ "Sex",
  pi_age ~ "Chronological age",
  pi_type ~ "Type of birth",
  oh_delivery ~ "Mode of delivery",
  outcome ~ "Death"
)

dat <- dat %>%
  mutate(oh_delivery = fct_recode(
    oh_delivery,
    instrumental = "forceps",
    instrumental = "ventouse"
  ))

dat %>%
  mutate(outcome = if_else(outcome == "died", 1, 0)) %>%
  select(all_of(t1_vars)) %>%
  tbl_summary(
    by = sepsis,
    statistic = list(
      c(pi_gest, et_bw) ~ "{mean} ({sd})",
      adm_dur ~ "{median} [{p25}-{p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = labels,
    missing = "no"
  ) %>%
  add_p(
    test = list(
      c(pi_gest, et_bw) ~ "t.test",
      adm_dur ~ "wilcox.test",
      c(pi_age, outcome) ~ "chisq.test.no.correct",
      c(pi_sex, pi_type, oh_delivery) ~ "fisher.test"
    )
  ) %>%
  add_overall() %>%
  add_stat_label(
    label = list(
      pi_gest ~ "mean weeks (SD)",
      et_bw ~ "mean grams (SD)",
      adm_dur ~ "median days [Q1-Q3]",
      all_categorical() ~ "n (%)"
    )
  ) %>%
  modify_footnote(update = everything() ~ NA) %>%
  as_kable_extra() %>% # to allow table to print in PDF output
  add_footnote(
    "p-values are from Welch’s two-sample t-test for gestational age and birth
    weight; the Wilcoxon-Mann-Whitney U test for admission duration; Pearson’s
    chi-squared test for age at admission and death; and Fisher’s exact test for
    sex, type of birth and mode of delivery.",
    notation = "none",
    threeparttable = TRUE
  )
```

<br />

Data are presented for the observed data only (i.e. before MICE) using pairwise
deletion of missing values. The number of missing values for each variable in
the above table are as follows:

```{r table1_missing}
dat %>% select(all_of(t1_vars)) %>% miss_var_summary()
```

## Distribution of candidate predictors

```{r predictors}
predictors <- c(
  "pi_gest",
  "et_bw",
  "oh_matfever",
  "oh_offliquor",
  "co_prom",
  "et_grunt",
  "et_cyanosis",
  "et_seizures",
  "et_rr",
  "et_hr",
  "et_temp",
  "oe_fontanelle",
  "oe_activity",
  "oe_nasalflare",
  "oe_retractions",
  "oe_grunt",
  "oe_wob",
  "oe_colour",
  "oe_abdodist",
  "oe_omphalitis",
  "oe_abskin",
  "hx_vomit"
)


dat %>%
  select(all_of(predictors), sepsis) %>%
  tbl_summary(
    by = sepsis,
    statistic = list(all_continuous() ~ "{median} [{p25}-{p75}]",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
    ) %>%
  add_p(
    test = list(all_continuous() ~ "wilcox.test",
                all_categorical() ~ "fisher.test")
  ) %>%
  add_overall() %>%
  as_gt(include = -tab_footnote) %>%
  tab_footnote(
    footnote = "Data are presented as median [Q1-Q3] for continuous predictors
    or n (%) for categorical predictors. p-values are from the Wilcoxon-Mann-Whitney
    U test for continuous predictors and Fisher’s exact test for categorical
    predictors.",
    locations = cells_column_labels(columns = vars(p.value))
  )
```

<br />

Data are presented for the observed data only (i.e. before MICE) using pairwise
deletion of missing values. The number of missing values for each variable in
the above table are as follows:

```{r predictors_missing}
dat %>% select(all_of(predictors)) %>% miss_var_summary()
```

### Box plots of continuous candidate predictors

The box plots below show the distribution of the continuous candidate predictors
between participants with and without sepsis.

```{r predictors_boxplots}
predictors_num <- c(
  "pi_gest",
  "et_bw",
  "et_rr",
  "et_hr",
  "et_temp"
  )

dat %>%
  select(all_of(predictors_num)) %>%
  map2(predictors_num, ~(
    ggplot(data = dat, aes(x = dat$sepsis, y = ., fill = dat$sepsis)) +
      geom_boxplot(colour = "black") +
      scale_fill_brewer() +
      labs(y = .y, x = "sepsis") +
      coord_flip() +
      theme_classic2() +
      theme(legend.position = "none") +
      NULL
    )
)
```

