---
title: "9-Reviewers"
output:
  pdf_document:
    number_sections: yes
    df_print: tibble
    latex_engine: xelatex
---

```{r setup9, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo=FALSE, message=FALSE, out.width="60%")
# Set root directory from Rmd dir to project dir
opts_knit$set(root.dir = "..") 
```

```{r libraries9, include=FALSE}
source("Scripts/0-setup.R")
```

```{r data9, include=FALSE}
incl_adm <- readRDS("Data/prepare2_dat.rds") %>% pull(adm_session)
incl_dis <- readRDS("Data/prepare2_dat.rds") %>% pull(dis_session)

adm <- readRDS("Data/prepare1_adm.rds") %>%
  filter(Admission.session %in% incl_adm)
dis <- readRDS("Data/prepare1_dis.rds") %>%
  filter(Discharge.session %in% incl_dis)
```

Additional analyses in response to reviewers' comments.

## Source of admission

The below figure shows the source of admission for neonates included in our study.

```{r adm_source}
adm_source <- adm %>% select(Admission.AdmittedFrom)
adm_tbl <- table(adm_source)
# proportions(adm_tbl)*100 # excludes NA data
# adm_tbl
miss_n <- sum(is.na(adm_source))
miss_p <- sum(is.na(adm_source))/nrow(adm_source)*100

adm_source_plot <- adm_source %>% 
  mutate(Admission.AdmittedFrom = fct_recode(
    Admission.AdmittedFrom,
    "admissions" = "ADM",
    "antenatal ward" = "ANC",
    "labour ward" = "LW",
    "postnatal ward" = "PNW",
    "theatre" = "TH"
    )) %>%
  mutate(Admission.AdmittedFrom = reorder(
    Admission.AdmittedFrom,
    Admission.AdmittedFrom,
    function(x) -length(x)
    )) %>%
  filter(!is.na(Admission.AdmittedFrom)) %>%
  ggplot(aes(x = Admission.AdmittedFrom)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, .6)) +
  labs(
    x = "Source of admission",
    y = "Proportion",
    caption = paste0(
      "*Excluding missing data (n=",
      miss_n, ", ", round(miss_p, 0), "%)")
    ) +
  theme_classic2() +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1, vjust = 1)) +
  NULL
adm_source_plot # excludes NA data
```

## Antibiotic usage

The below figure shows the distribution of antibiotic usage for neonates
included in our study.

```{r abx_usage}
# Available antibiotics as per data dictionary
abx <- list("BP", "GENT", "AMOX", "AMIK", "AMP", "CEF", "CLOX", "IMI",
            "MET", "PROC", "VAN") # excludes chloramphenicol
# sum(is.na(dis$Discharge.MedsGiven)) #=0

abx_usage <- dis %>%
  select(Discharge.session, Discharge.MedsGiven) %>%
  mutate(
    benzypenicillin = if_else(grepl("BP", Discharge.MedsGiven), 1, 0),
    gentamicin = if_else(grepl("GENT", Discharge.MedsGiven), 1, 0),
    amoxicillin = if_else(grepl("AMOX", Discharge.MedsGiven), 1, 0),
    amikacin = if_else(grepl("AMIK", Discharge.MedsGiven), 1, 0),
    ampicillin = if_else(grepl("AMP", Discharge.MedsGiven), 1, 0),
    ceftrixone = if_else(grepl("CEF", Discharge.MedsGiven), 1, 0),
    cloxicallin = if_else(grepl("CLOX", Discharge.MedsGiven), 1, 0),
    imipenem = if_else(grepl("IMI", Discharge.MedsGiven), 1, 0),
    metronidazole = if_else(grepl("MET", Discharge.MedsGiven), 1, 0),
    `procaine pencillin` = if_else(grepl("PROC", Discharge.MedsGiven), 1, 0),
    vancomycin = if_else(grepl("VAN", Discharge.MedsGiven), 1, 0)
  ) %>%
  select(-Discharge.MedsGiven)

abx_tbl <- abx_usage %>%
  mutate(across(-Discharge.session, ~ sum(.))) %>%
  select(-Discharge.session) %>%
  unique() %>%
  pivot_longer(everything())
# abx_tbl %>% arrange(desc(value))

abx_plot <- abx_tbl %>%
  ggplot(aes(x = fct_rev(fct_reorder(name, value)), y = value)) +
  geom_col() +
  labs(
    x = "Antibiotic",
    y = "Number of neonates") +
  theme_classic2() +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1, vjust = 1)) +
  NULL
abx_plot
```

Below is the distribution of antibiotic usage for those with and without EOS.

```{r abx_usage2}
# sum(rowSums(abx_usage == 1) > 0) # received abx during admission
# sum(rowSums(abx_usage == 1) > 0)/nrow(dis)*100 # = percentage prescribed abx

# Chi-sq test for abx use x sepsis to include in Table 1
a <- abx_usage %>% mutate(abx = if_else(rowSums(. == 1) > 0, 1, 0)) %>%
  select(Discharge.session, abx) %>%
  rename(dis_session = Discharge.session)
b <- readRDS("Data/prepare2_dat.rds") %>% select(dis_session, sepsis)
abx_dat <- merge(a, b, by = "dis_session")
# nrow(abx_dat)

table("abx (n)" = abx_dat$abx, "sepsis (n)" = abx_dat$sepsis)
round(prop.table(table("abx (%)" = abx_dat$abx, "sepsis (%)" = abx_dat$sepsis), margin = 2)*100, 0)
chisq.test(table(abx_dat$abx, abx_dat$sepsis))
```


## Characteristics of neonates in whom sepsis 'missed' by model

The table below compares characteristics between those with EOS who were classified
as EOS by our model (true positives) and those with EOS who were classified as 'no EOS'
by our model (false negatives, i.e. 'missed' EOS) at the optimal classification threshold
determined by Youden's _J_ statistic (0.121).

```{r sepsis_missed}
si <- readRDS("Data/si.rds")
pred_si <- readRDS("Data/pred_si.rds") # predictions from model
thresopt <- 0.121

si <- si %>%
  mutate(sepsis_pred = if_else(pred_si > thresopt, "yes", "no")) %>%
  mutate(missed = if_else(sepsis == "yes" & sepsis_pred == "no", 1, 0))
# table(si$missed)

# should be same order in both df (excluding missing values)
missed_dat <- readRDS("Data/prepare2_dat.rds") %>%
  mutate(et_bw = et_bw/1000,
         oh_delivery = fct_recode(
           oh_delivery,
           instrumental = "forceps",
           instrumental = "ventouse"
))
# just to make sure
# missed_dat$et_bw[!mapply(identical, si$et_bw, missed_dat$et_bw)]
# only the missing values = this is OK

missed_dat <- missed_dat %>% mutate(missed = si$missed)

t1_vars <- c(
  "pi_sex",
  "pi_gest",
  "et_bw",
  "pi_age",
  "pi_type",
  "oh_delivery",
  "adm_dur",
  "outcome",
  "sepsis",
  "missed"
)

missed_labels <- list(
  pi_gest ~ "Gestational age",
  et_bw ~ "Birth weight",
  adm_dur ~ "Admission duration",
  pi_sex ~ "Sex",
  pi_age ~ "Chronological age",
  pi_type ~ "Type of birth",
  oh_delivery ~ "Mode of delivery",
  outcome ~ "Death"
)

missed_dat %>%
  mutate(outcome = if_else(outcome == "died", 1, 0)) %>%
  select(all_of(t1_vars)) %>%
  filter(sepsis == "yes") %>% # just compare those with sepsis who were missed vs. not missed
  tbl_summary(
    by = missed,
    statistic = list(
      c(pi_gest, et_bw) ~ "{mean} ({sd})",
      adm_dur ~ "{median} [{p25}-{p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = missed_labels,
    missing = "no"
  ) %>%
  add_p(
    test = list(
      c(pi_gest, et_bw) ~ "t.test",
      adm_dur ~ "wilcox.test",
      all_categorical() ~ "fisher.test"
    )
  ) %>%
  add_overall() %>%
  add_stat_label(
    label = list(
      pi_gest ~ "mean weeks (SD)",
      et_bw ~ "mean grams (SD)",
      adm_dur ~ "median days [Q1-Q3]",
      all_categorical() ~ "n (%)"
    )
  ) %>%
  modify_footnote(update = everything() ~ NA) %>%
  as_kable_extra() # %>% # to allow table to print in PDF output

```


## Reasons for admission

The below figure shows the reasons for admission of neonates included in our study.

```{r adm _reason}
# R2.5 - Reasons for admission --------------------

adm_reason <- adm %>%
  select(Admission.AdmReason) %>%
  mutate(
    Admission.AdmReason = fct_recode(
      Admission.AdmReason,
      "abdominal distention" = "AD",
      "low apgars" = "Apg",
      "HIE" = "BA",
      "born before arrival" = "BBA",
      "congenital anomaly" = "Cong",
      "convulsions" = "Conv",
      "resp distress" = "DIB",
      "abandoned" = "DU",
      "feeding difficulty" = "FD",
      "fever" = "Fev",
      "gastroschisis" = "G",
      "HIV exposed" = "HIVX",
      "jaundice" = "J",
      "low birth weight" = "LBW",
      "macrosomia" = "Mac",
      "meconium" = "Mec",
      "neural tube defect" = "NTD",
      "other" = "O",
      "omphalocele" = "OM",
      "prematurity" = "Prem",
      "prematurity with resp distress" = "PremRDS",
      "risk of sepsis" = "Risk",
      "safekeeping" = "Safe",
  )
)

# table(adm_reason$Admission.AdmReason)

adm_reason_plot <- adm_reason %>%
  mutate(Admission.AdmReason = reorder(
    Admission.AdmReason,
    Admission.AdmReason,
    function(x) length(x)
  )) %>%
  ggplot(aes(x = Admission.AdmReason)) +
  geom_bar() +
  coord_flip() +
  labs(
    x = "Admission reason",
    y = "Number of neonates") +
  theme_classic2() +
  NULL
adm_reason_plot
```


## Sample size calculations

The below code shows post hoc sample size calculations as per Riley _et al._ BMJ 2020
(doi: https://doi.org/10.1136/bmj.m441).

```{r sample_size, echo=TRUE}
# a. estimate the overall outcome proportion with sufficient precision
# aiming for margin of error \leq 0.05
outprop <- sum(si$sepsis == "yes")/nrow(si) #outcome proportion
(1.96 / 0.05)^2 * outprop*(1-outprop) #=154

# b. target a small mean absolute prediction error (p \leq 30)
n <- nrow(si) #sample size of development dataset
P <- 9 #number of candidate predictor parameters
MAPE <- 0.05 #mean absolute prediction error
exp((-0.508 + 0.259*log(outprop) + 0.504*log(P) - log(MAPE))/0.544) #=263 for MAPE 0.05

# c. target a shrinkage factor of 0.9
# (although shrinkage not used in our study)
S <- 0.9 #target shrinkage factor of 10%
R2CS <- 0.1 #anticipated R2(CS) of at least 0.1
P / ((S-1) * log(1 - (R2CS/S))) #=764

# d. target small optimism of 0.05 in the apparent R2(Nagelkerke)
lnLnull <- (outprop*n)*log(outprop*n/n) + (n - outprop*n)*log(1 - outprop*n/n)
maxR2CS <- 1 - exp(2*lnLnull / n) #maximum possible value of R2CS
S2 <- R2CS / (R2CS + 0.05*maxR2CS) #assuming expected optimism \leq 0.05
P / ((S2 - 1)*log(1 - R2CS/S2)) #=333
```


## Model calibration

The below statistics and figure shows calibration of our optimal model with a
flexible calibration curve using a loess smoother.

```{r calibration, warning=FALSE}
si <- readRDS("Data/si.rds")
pred_si <- readRDS("Data/pred_si.rds")
cal_stats <- CalibrationCurves::valProbggplot(
  pred_si, if_else(si$sepsis == "yes", 1, 0),
  dostats = c("Intercept", "Slope"), #simplifies label of intercept/slope
  col.smooth = "red",
  col.ideal = "black",
  lwd.ideal = 0.5,
  size = 4
)
cal_stats$stats

cal_plot<- cal_stats$ggPlot +
  theme_classic2() +
  theme(legend.position = "top") +
  NULL

cal_violin <- ggplot(si, aes(x = if_else(sepsis == "yes", 1, 0), y = !!pred_si, fill = sepsis)) +
  geom_violin(colour = "black") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0,1,.25)) +
  scale_x_continuous(breaks = c(0, 1)) +
  scale_fill_manual(values = c("grey", hcl(0, 100, 40)),
                    guide = guide_legend(reverse = T)) +
  new_scale_fill() +
  geom_boxplot(fill = "white", width = .15, outlier.shape = 1) +
  labs(y = "Predicted probability of sepsis", x = "Observed case of sepsis",
       fill = "Observed sepsis: ") +
  coord_flip() +
  annotate("text", x = 0, y = .9, label = "0") + 
  annotate("text", x = 1, y = .9, label = "1") + 
  theme_classic2() +
  theme(axis.ticks=element_blank(),
        axis.title=element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        plot.margin = margin(5.5,5.5,5.5,47, "pt")) +
  NULL

cal_fig <- plot_grid(cal_plot, cal_violin, nrow = 2, rel_heights = c(4,1))
cal_fig

ggsave("Figures/calibration_fig.png", cal_fig, width = 5, height = 5, dpi = 300)
```

