---
title: "Supplementary Methods & Results"
subtitle: "Diagnosing early-onset neonatal sepsis in low-resource settings: development of a multivariable prediction model"
author: "Neal SR et al."
date: "02/03/2023"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 1
    
    df_print: tibble
    highlight: tango
fontsize: 10pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{afterpage}
---

\vspace{30pt}

Some supplementary data have been redacted or replaced by fictitious examples (where indicated) to remove identifying information for publication. 

\newpage

# Neotree data collection

```{r child="1-NeotreeDataCollection.Rmd"}
```

\newpage

# Candidate predictors

```{r child="2-CandidatePredictors.Rmd"}
```

\newpage

# Preliminary data cleaning

```{r child="3-PreliminaryCleaning.Rmd"}
```

\newpage

# Record linkage

```{r child="4-RecordLinkage.Rmd"}
```

\newpage

# Further data cleaning

```{r child="5-FurtherCleaning.Rmd"}
```

\newpage

# Missing data

```{r child="6-MissingData.Rmd"}
```

\newpage

# Descriptive statistics

```{r child="7-DescriptiveStatistics.Rmd"}
```

\newpage

# Model development and performance

```{r child="8-Modelling.Rmd"}
```

\newpage

# Additional analyses for reviewers' comments

```{r child="9-Reviewers.Rmd"}
```
